Marginalization of photometric redshift uncertainties

### 2020-07-04 Paper/paper.tex
Updates from Overleaf

\documentclass[a4paper,11pt]{article}
\usepackage{ifpdf} 
\ifpdf
\pdfoutput=1 % if your are submitting a pdflatex (i.e. if you have
             % images in pdf, png or jpg format)
\fi

\usepackage{jcappub}
\usepackage{graphicx}
\graphicspath{{Paper/}{./}}
\usepackage{dcolumn}
\usepackage{amssymb,amsmath,bm}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
%\usepackage[utf8]{inputenc}
\usepackage{xfrac}
\usepackage{aas_macros}
\usepackage{mathrsfs}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{chngcntr}
\newcommand{\nv}{\vec{\theta}}
%\newcommand{\pasj}{Publications of the ASJ}
%\newcommand{\aap}{Astronomy and Astrophysics}
%\newcommand{\apj}{Astrophysical Journal}
%\newcommand{\apjs}{Astrophysical Journal, Supplement}
%\newcommand{\mnras}{Monthly Notices of the RAS}
\newcommand{\todo}[1]{{\bf TODO: #1}}

@@

%\usepackage{lineno}
%\linenumbers

@@

\section{Introduction}\label{sec:intro}

Photometric galaxy surveys  have the potential of  being transformative in our understanding of the Universe by measuring millions and soon billions of galaxies on the sky. These catalogs can be used to constrain cosmological parameters through measurements of clustering and weak lensing as demonstrated by numerous surveys that have completed observations including  Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201}.

@@

\section{Theory}

In typical cosmological analysis problem, we are performing a Bayesian inference on the posterior $p({\rm theory} | {\rm data} ) \propto p({\rm data}|{\rm} theory) p({\rm theory})$ with the data likelihood given by 
\begin{equation}
p({\rm data}|{\rm theory}) =  \mathcal{L} \propto  \exp\left [-\frac{1}{2} \chi^2 \right], \label{eq:like}
\end{equation}
where
\begin{equation}
  \chi^2  = (\vd-\vt)^T C^{-1} (\vd-\vt),
\end{equation}
where $\vd$ is the data vectory, typically containing all the summary statistic of interest like correlation functions ot power spectra and $\vt$ the theory vector through which parameter dependence enters. In this set-up the covariance matrix $C$ is fixed (and hence the Gaussian normalization term does not enter equation \ref{eq:like}), which is usually a very good approximation in typical problems that we wish to consider \cite{1811.11584}, but note that the method is easily generalizable to the cases where this is not true.

We can make significant progress without specifying explicitly how theory is calculate given the parameters. It suffices to say that given a set of cosmological parameters, which specify the evolution of the universe and thus matter power spectra as a function of scale and redshift, a set of survey properties, such as number densities and redshift distribution of galaxies and weak lensing sources, we can integrate these into theory predictions $\vt=\vt(\theta)$.

The traditional approach is to add to add nuissance parameters that describe our lack of knowledge of galaxy redshift distribution. A typical approach is to use shift and width parameters  \cite{1706.09359,1912.08209} \as{did DES Y1 have shift/width?},
where redshift distributions $N_i(z)$ are shifted according to
    \begin{equation}
      N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right),
      \label{eq:photo-z-model}
    \end{equation} 
where the index $i$ runs over the number of tomographic samples. Here $N_{i}(z)$ is the true redshift distribution with  $\bar{N}_{i}(z)$ out best estimate of it. The parameters $\Delta z_{i}$ and $z_{w,i}$ account for shifts in the means of the distributions, while $z_{c,i} = \left<z\right>_{\bar{N}_i}$ is the mean redshift to act as a sensible pivot for width changes. The shift and width parameters for each tomographic bin are then inferred and marginalized over together with all other parameters.

This approach explicitly marginalizes over the two effects that we think are the two most important systematics stemming from redshift distribution uncertainty. However, it suffers from two problems. First, it is fundamentally ad-hoc. This particular parametric form does not stem from consideration of phometric redshift estimators. There is also no reason to believe why it would be complete in the sense that it captures all relevant effects. The second problem is that it is computationally expensive. For example, in the analysis 
    of \cite{1912.08209}, redshift uncertainty parameters accounted for 8 (2 per 4 tomographic bins) of 14 total parameters in the inference process, i.e. over the half parameters inferred were dealing with redshift.

\subsection{A new approach}

In this paper we advocate a different approach. First, let's consider a binned description of $N_i(z)$ in terms of $N_{ij}$, where $i$ index describes the tomographic bin and $j$ is redshift index, so that $N_{ij} = N_i(z_j)$ at some discrete grid of $z_j$. This entire vector can be packed in a vector $\vN$.

We will describe our uncertainty in $\vN$ as a Gaussian
\begin{equation}
  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
\end{equation}
where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations. We will discuss later how $P$ can be determined, but at this point we note that it is a general positive-definite matrix, i.e not necessarily diagonal). Other possible parameterizations are sub-summed in this approach. For example, a general shift-width is allowed with the probability $p(\vN(\Delta z_i, z_{w,i}))$.

We can now write a likelihood for the data in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
\begin{equation}
\mathcal{L} \propto  \int \exp\left [-\frac{1}{2} (\vd-\vt)^T C^{-1} (\vd-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
  \end{equation}
  Note that in the equation above, $\vt=\vt(\theta,\vN)$, where $\theta$ are the remaining parameters. It is this dependence that prevents us from performing what looks like a trivial integral.

  In principle, we could solve this problem by making every single element of vector $\vN$ a part of MCMC chain and simply run the MCMC with $\theta$ and additional hundreds of $\vN$ parameters. Instead, we pull an Eastern-European trick and Taylor expand theory around $\hat{\vN}$:
  \begin{equation}
    \vt(\theta,\vN) = \vt(\theta,\bar{\vN}) + T \left(\vN - \bar{\vN} \right), \label{eq:taylor}
  \end{equation}
  where
  \begin{equation}
    T_{mn} = \frac{\partial t_m}{\partial N_n} \bigg|_{\theta,\hat{\vN}}
  \end{equation}
  The matrix $T$ is the non-square matrix giving derivatives of theory with respect to each element of $\vN$. In other words it contains a full information about how all the correlation functions or power spectra respond to changing $N(z)$ distribution for one tomographic bin at one redshift.

  Substituting Eq. \ref{eq:taylor} into Eq. \ref{eq:like} results in an integral that is now quadratic in $\bar{\vN}$ and can be performed analytically. After collecting terms that are quadratic and linear in $\vN$, performing the Gaussian integral and applying the Woodbury matrix identity, we find:
  \begin{equation}
    \mathcal{L} \propto \frac{1}{|T C^{-1} T^T +P^{-1} |^{1/2}} \exp\left [-\frac{1}{2} (\vd-\vt)^T C_M^{-1} (\vd-\vt) \right], \label{eq:main1}
  \end{equation}
  where
  \begin{equation}
    C_M = C + T PT^T \label{eq:main2}
  \end{equation}
  This is a fascinatingly simple equation and the main result of this paper. This calculation should be understood as follows: for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$. These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.  Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far P allows them to go.

  In principle, at every point in likelihood, one should evaluate Eq. \ref{eq:main1}, which means recalculating $T$. Since this is a very expensive computation, we will further simplify the problem, by assuming $T$ is only weakly deviations from some fiducial cosmology. We will therefore pre-compute $T$ once for the fiducial cosmology, which then means that normalization term in equation \ref{eq:main1} is an irrelevant constant and at fixed data covariance matrix $C$, the Eq. \ref{eq:main2} also needs to be evaluated just once.

  The procedure is thus as follows:
  \begin{itemize}
  \item Run inference problem once with fixed redshift distribution to determine a sensible fiducial model.

  \item Calculate $T$ and $C_M$ at this fiducial model

  \item Re-run inference with the modified covariance matrix $C_M$, which leads to broadened contours due to marginalization of redshift distributions

  \end{itemize}
If needed, one could repeat the inference at a refined fiducial model although in pracice we found this to be unnecessary.

Note that there are two distinct approximation here. The first is that theory can validly be Taylor expanded in $\vN-\bar{\vN}$ and the second is that $T$ is model independent for models of interest. We will examine validity of both of those approximations in Section \ref{sec:hsc}.

\subsection{Determination of prior matrix $P$}

Determination of the prior matrix $P$ is where the the actual and often partially subjective prior on our knowledge of $N(z)$ enters. Note that the standard parameterizations can always be recast in this language at least in the perturbative limits. For example, a vector of derivatives $\lambda_i = dN_i/dz$ when added to $\vN$ will produce a shift in the the shape of $N_i(z)$. Therefore, adding a contribution $\sigma_{zi}^2 \lambda_i \lambda_i^T$ will allow shifts with variance $\sigma_{zi}$. Of course, when shifts become so large that the perturbative expansion breaks, the approach will work only approximately. But in this paper we are not interesting in shifts; instead we will try to build $P$ is a more systematic manner.

\subsubsection{Uncertainty due to estimator biases}
We will build $P$ from three pieces. The first piece is simply the overall uncertainty stemming from the fact that the photometric redshift estimators have errors. The simplest thing is to make this covariance matrix   \ldots

\subsubsection{Cosmic Variance in calibrator sample}
The covariance matrix between two tomographic bins is given by:
\begin{equation}
Cov_{ij}=\frac{N_i N_j}{2\pi^2} \int_0^\infty dk_\parallel \cos(k_\parallel(\chi_i-\chi_j))\int_0^\infty dk_\perp k_\perp W_i(k_\parallel,k_\perp)W_j(k_\parallel,k_\perp)P_{\bf gg}({\bf k}),
\end{equation}
where we can model the power spectra at each bin centre and for all
transverse and parallel modes using the Kaiser formula (CITE), which corrects for redshift space distortions through
the expression:
\begin{equation}
P_{gg}({\bf k},z)=\left(b_g(z)+f(z)\frac{k_\parallel^2}{k^2}\right)^2\,P_{mm}(k,z),
\end{equation}
where $f(z)$ is
the logarithmic growth
rate
and the
evolution of the linear galaxy bias is given by $b_g(z) = 0.95/D(z)$
with $D(z)$ the growth factor.

In this work, we use the galaxy sample in the COSMOS survey as calibrator, which
covers a total area of
$A_{\rm sky} = 1.7 \ {\rm deg}^2$ for galaxies with redshifts $z < 2$.
We model the window function for each bin of the dataset through:
\begin{equation}
W(k_\parallel,k_\perp)=j_0(k_\parallel\Delta\chi/2)\,\frac{2 J_1(k_\perp R_{\rm mean})}{k_\perp R_{\rm mean}},
\end{equation}
where $R_{mean}$ is the
physical size of the survey measured at the mean
distance to the
tomographic bin, 
$R_{\rm mean} = 
\sqrt{\frac{A_{\rm sky}}{\pi}} \chi_{\rm mean}$.
We assume that the
cosmic covariance
between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by
treating each bin independently. We show the resulting correlation matrix of the COSMOS cosmic variance in Fig. \ref{fig:CV}.

\begin{figure}[ht]
\centering  
\includegraphics[width=1.\textwidth]{./corr_CV_0.png}
\caption{Cosmic variance (CV) correlation matrix with subtracted diagonal for the first tomographic bin of the HSC data.} 
\label{fig:CV}
\end{figure}

\subsubsection{Smoothness prior}

\section{Appplication to HSC clustering data}
\label{sec:hsc}
\an{Some citations are currently missing but I will put them in later.}
We apply the method outlined above to the data obtained in Ref.~\cite{1912.08209}. This work measured the angular galaxy clustering power spectrum from HSC DR1 data. In the following, we give a very brief summary of the methodology employed in this work and refer the reader to Ref.~\cite{1912.08209} for more details.
\subsection{Background theory}
The angular clustering signal for galaxies in redshift bins $i, j$ with photo-$z$ distributions $p^i(z), p^j(z)$ can be modeled using the Limber approximation as \cite{1953ApJ...117..134L, 1992ApJ...388..272K, Kaiser:1998}
    \begin{equation}\label{eq:cell_gg_limber}
      C^{ij}_\ell = \int \mathrm{d}z\,\frac{H(z)}{\chi^2(z)} p^i(z)p^j(z)\,P_{gg}\left(z,k=\frac{\ell+1/2}{\chi(z)}\right),
    \end{equation}
where $P_{gg}(z,k)$ denotes the underlying 3D galaxy power spectrum, $\chi(z)$ is the comoving distance and $H(z)$ denotes the Hubble parameter at redshift $z$. As can be seen from Eq.~\ref{eq:cell_gg_limber}, the spherical harmonic power spectrum $C_{\ell}$ depends linearly on the photometric redshift distributions $p^i(z)$, which facilitates the computation of the $T$ matrix defined above.

We follow Ref.~\cite{1912.08209} and compute theoretical predictions for the galaxy power spectrum $P_{gg}(z,k)$ within the halo model combined with halo occupation distribution modeling \cite{2000MNRAS.318.1144P,2002PhR...372....1C,2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}. 
\subsection{The HSC dataset}
The Hyper-Suprime Cam survey is an on-going photometric galaxy survey survey focused mainly on weak gravitational lensing. The analysis in Ref.~\cite{1912.08209} is based on the publicly-available HSC DR1 data, whose so-called wide fields cover approximately 108 square degrees on the sky, subdivided into seven distinct patches. Ref.~\cite{1912.08209} used these data to compute spherical harmonic galaxy clustering power spectra for four tomographic redshift bins between $z=0.15$ and $z=1.5$, taking both auto- and cross-correlations into account. These power spectra have been corrected for observational and extragalactic systematics by deprojection at the map-level \cite{2019MNRAS.484.4127A}. Finally, following Ref.~\cite{2019PASJ...71...43H}, photometric redshift distributions have been estimated by cross-matching HSC galaxies to galaxies in the COSMOS 30-band photometric catalog presented in Ref.~\cite{2016ApJS..224...24L}.
\subsection{Validating the linear expansion}

>>>>

\documentclass[a4paper,11pt]{article}
\usepackage{ifpdf} 
\ifpdf
\pdfoutput=1 
\fi

\usepackage{jcappub}
\usepackage{graphicx}
\graphicspath{{Paper/}{./}}
\usepackage{dcolumn}
\usepackage{amssymb,amsmath,bm,bbold}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{xfrac}
\usepackage{aas_macros}
\usepackage{mathrsfs}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{chngcntr}
\newcommand{\nv}{\vec{\theta}}
\newcommand{\todo}[1]{{\bf TODO: #1}}

@@

%\linenumbers

@@

  \section{Introduction}\label{sec:intro}\label{sec:intro}
    Photometric galaxy surveys  have the potential of  being transformative in our understanding of the Universe by measuring millions and soon billions of galaxies on the sky. These catalogs can be used to constrain cosmological parameters through measurements of clustering and weak lensing as demonstrated by numerous surveys that have completed observations including  Sloan Digital Sky Survey \cite{astro-ph/0006396}, PanSTARRS \cite{2010SPIE.7733E..0EK}, Kilo-degree Survey \cite{1206.1254}, Dark Energy Survey \cite{1601.00329}, Hyper Suprime-Cam survey \cite{2012SPIE.8446E..0ZM}, and will play a crucial role in the upcoming Vera Rubin Observatory LSST Survey \cite{0912.0201}.

@@

  \section{Theory}\label{sec:theory}
    \subsection{Redshift distribution uncertainties}\label{ssec:theory.nz}
      Most cosmological analyses involve performing Bayesian parameter inference on a posterior distribution $p(\vec{\theta} |\vd) \propto p(\vd|\vec{\theta}) p(\vec{\theta})$, where $\vd$ is a vector of data points and $\vec{\theta}$ is a set of parameters descring the underlying model.

      In the case of tomographic analyses of galaxy clustering and large-scale structure, $\vd$ traditionally contains correlation function or power spectrum measurements between different tracers (e.g. galaxy ellipticities or overdensity) and different redshift bins\footnote{We use the term ``redshift bin'' here to denote any given galaxy sample, whether or not it has actually been selected by binning those galaxies into intervals of some redshift estimate.}, and $\vec{\theta}$ includes both cosmological and nuisance parameters needed to produce a forward model of $\vd$ that we will call $\vt(\vec{\theta})$. Due to the central limit theorem it is often accurate enough to assume that the likelihood is Gaussian, taking the form
      \begin{equation}
        \mathcal{L} \equiv p({\bf d}|\vec{\theta}) = \frac{\exp\left [-\frac{1}{2}(\vd-\vt(\vec{\theta}))^T{\sf C}^{-1}(\vd-\vt(\vec{\theta})) \right]}{\sqrt{\det{2\pi{\sf C}}}}, \label{eq:like}
     \end{equation}
      where ${\sf C}$ is the covariance matrix of $\vd$. The information contained in the parameter dependence of ${\sf C}$ can be neglected \citep{1811.11584}, and therefore the normalization factor $\sqrt{2\pi\det{\sf C}}$ can be ignored.

      We can make significant progress without specifying explicitly how the theory is calculated given the parameters. It suffices to say that, given a set of cosmological parameters, which specify the evolution of the universe and the matter inhomogeneities as a function of scale and redshift, a set of astrophysical parameters (e.g. a particular bias or intrinsic alignment model), and the redshift distributions $N(z)$ of the different redshift bins, we can integrate these into theory predictions $\vt=\vt(\vec{\theta})$. Our main concern in this paper is the way in which the significant uncertainties in $N(z)$, which can potentially dominate the total systematic error budget, should be parametrized and marginalized over.

      The traditional approach (e.g. \todo{Add cites}) has been to parametrize departures from a given ``educated guess'' of the redshift distribution that are likely to describe the main modes in which the underlying uncertainty propagates into the theory vector. This has usually been done by introducing ``shift'' ($\Delta z$) and/or ``width'' ($z_w$) nuisance parameters, in terms of which the fiducial redshift distribution for the $i$-th redshift bin $\bar{N}_i(z)$ is modified as
      \begin{equation}
        N_{i}(z) \propto \bar{N}_{i}\left(z_{c,i} + (1 + z_{w, i})(z-z_{c, i}) + \Delta z_{i}\right), \label{eq:photo-z-model}
      \end{equation}
      where $z_{c,i}$ is the mean of the fiducial distribution and acts as a sensible pivot to define width changes. $\Delta z_i$ and $z_{w,i}$ for each bin are then inferred and marginalized over together with all other parameters, with priors on them based on calibration uncertainties.

      This approach explicitly marginalizes over two effects that are potentially the two most important systematics stemming from redshift distribution uncertainty: the width of the lens redshift distribution affects the amplitude and shape of the clustering power spectrum, and a shift in the distribution of sources influences the support of the lensing kernel and the amplitude of the shear two-point function. However, this methods suffers from two shortcomings. First, it is fundamentally ad-hoc: this particular parametric form is not directly mapped onto the main modes of uncertainty in photometric redshift estimators \todo{we may want an expert to vet this statement}, and there is rigorous proof that the parametrization is complete in the sense that it captures all relevant effects. The second problem is that it is computationally expensive. By adding one or two additional nuisance parameters per redshift bin, the dimensionality of the model parameter space increases significantly. For instance, in the analysis of \cite{1912.08209}, redshift uncertainty parameters accounted 50-60\% of all model parameters. The methodology described in the next section aims to address both of these shortcomings.

    \subsection{A new approach}\label{ssec:theory.nz_new}
      Let us consider a discretized description of the $N_i(z)$ as a sum over basis functions $\phi_\alpha(z)$:
      \begin{equation}
        N_i(z)=\sum_i N^\alpha_i\phi_\alpha(z).
      \end{equation}
      For simplicity here we will just treat $N_i(z)$ as a histogram, and therefore in our case the $\phi_\alpha$ are just top-hat functions centered on an equi-spaced grid of redshifts $z_\alpha$. The formalism described below is valid for any choice of basis functions, however. In this formalism, full set of redshift distributions is described by the vector $\vN\equiv\{N^\alpha_i\,\,\,\forall i,\alpha\}$. In what follows, we will assume that $\vN$ has units of ``counts''. The actual normalization of the $N(z)$ is irrelevant, since it only enters the theory prediction for the power spectrum as a normalized redshift probability distribution (see Eq. \todo{ref}).

      Let us further distinguish between the true underlying redshift distribution, represented by $\vN$, and a possible measurement of it $\hat{\bf N}$, with measurement uncertainties encapsulated by a covariance matrix $\hat{\sf C}_N$. Furthermore, we may have some external prior information on $\vN$, which for simplicity we will assume to be Gaussian with mean $\vN_P$ and covariance $\hat{\sf C}_P$.

      Our data is therefore made up of the combination ${\bf d}=(\hat{\bf c},\hat{\bf N})$, where $\hat{\bf c}$ is a set of two-point function measurements. The model parameters are $\vec{\theta}=({\bf q},\vN)$, where $\vN$ are the redsihft distribution coefficients, and ${\bf q}$ contains all other cosmological, astrophysical and nuisance parameters. The posterior distribution is therefore given by:
      \begin{align}
        p(\vec{\theta}|{\bf d})
        &=p({\bf q},\vN|\hat{\bf c},\hat{\bf N})\\\label{eq:posterior_full}
        &=p(\hat{\bf c}|{\bf q},\vN)p(\hat{\bf N}|\vN)p(\vN)p({\bf q})\\
        &=p(q)\frac{\exp\left[-\frac{1}{2}(\chi^2_c+\chi^2_N+\chi^2_P)\right]}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
      \end{align}
      where
      \begin{align}
        \chi^2_c&=\left(\hat{\bf c}-{\bf t}\right)^T{\sf C}_c^{-1}\left(\hat{\bf c}-{\bf t}\right)\\
        \chi^2_N&=\left(\hat{\bf N}-\vN\right)^T{\sf C}_N^{-1}\left(\hat{\bf N}-\vN\right)\\
        \chi^2_P&=\left(\vN-\vN_P\right)^T{\sf C}_P^{-1}\left(\vN-\vN_P\right),
      \end{align}
      and ${\bf t}(q,\vN)$ is the theoretical prediction for $\hat{\bf c}$.

      In Eq. \ref{eq:posterior_full} we have assumed that $\hat{\bf c}$ and $\hat{\vN}$ are independent at the likelihood level, and that the likelihood of $\hat{\vN}$ is independent of ${\bf q}$. The validity of these assumptions should be studied in detail, especially in cases where $\vN$ is calibrated using a spectroscopic sample with significant spatial overlap with the data under study (through a clustering redshifts approach \todo{cite} or otherwise). The method outlined here is still applicable if these assumptions are dropped, albeit with a different expression for the modified covariance in Eq. \ref{eq:main2}. We have also assumed that the redshift distribution amplitudes $\hat{\vN}$ are Gaussianly distributed. \todo{add some blab here and potentially cite Carles Sanchez's latest paper.}

      To derive the main result of this paper, let us start by considering the combination $\chi^2_N+\chi^2_P$. After completing squares for $\vN$, this can be written as:
      \begin{equation}
        \chi^2_N+\chi^2_P\equiv\chi^2_{\bar{N}}=\left(\vN-\tilde{\vN}\right)^T{\sf P}^{-1}\left(\vN-\tilde{\vN}\right)+K_N,
      \end{equation}
      where we have defined the \emph{smoothed mean} $\bar{\vN}$ and combined prior covariance ${\sf P}$ as:
      \begin{align}\label{eq:priorcov}
        \bar{\vN}\equiv{\sf P}({\sf C}_N^{-1}\hat{\bf N}+{\sf C}_P^{-1}\vN_P),\hspace{12pt} {\sf P}^{-1}\equiv{\sf C}_N^{-1}+{\sf C}_P^{-1},
      \end{align}
      and $K_N$ is independent of the model parameters, and given by
      \begin{equation}
        K_N\equiv\hat{\bf N}^T{\sf C}^{-1}_N\hat{\bf N}+\vN_P^T{\sf C}_P^{-1}\vN_P-\bar{\vN}^T{\sf P}^{-1}\bar{\vN}.
      \end{equation}
      The full posterior thus takes the form:
      \begin{align}
        p({\bf q},\vN|\hat{\bf c},\hat{\bf N})&=Q\,p({\bf q})\exp\left[-\frac{1}{2}(\hat{\bf c}-{\bf t})^T{\sf C}^{-1}_c(\hat{\bf c}-{\bf t})-\frac{1}{2}(\vN-\bar{\vN})^T{\sf P}^{-1}(\vN-\bar{\vN})\right],\\
        Q&\equiv\frac{e^{-K_N/2}}{\sqrt{{\rm det}(2\pi{\sf C}_c){\rm det}(2\pi{\sf C}_N){\rm det}(2\pi{\sf C}_P)}},
      \end{align}

%We will describe our uncertainty in $\vN$ as a Gaussian
%\begin{equation}
%  p(\vN) \propto \exp\left[  (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right],
%\end{equation}
%where $\hat{\vN}$ is our best guess of redshift distribution $P$ is a matrix of allowed deviations. We will discuss later how $P$ can be determined, but at this point we note that it is a general positive-definite matrix, i.e not necessarily diagonal). Other possible parameterizations are sub-summed in this approach. For example, a general shift-width is allowed with the probability $p(\vN(\Delta z_i, z_{w,i}))$.

      We can now write a likelihood for $\hat{\bf c}$ in which we explicitly marginalize out all the degrees of freedom associated with $\vN$:
      \begin{equation}
        \mathcal{L} \propto  \int \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T C^{-1} ({\bf c}-\vt) -\frac{1}{2} (\vN - \bar{\vN})^T P^{-1} (\vN -\bar{\vN}) \right]  d^N\vN, \label{eq:like2}
      \end{equation}
      Note that in the equation above, $\vt=\vt({\bf q},\vN)$. It is the non-linear dependence of $\vt$ on $\vN$ that prevents us from performing what looks like a trivial integral analytically. In principle, we could solve this problem by making every single element of vector $\vN$ a part of an MCMC chain for ${\bf q}$ and additional hundreds of $\vN$ parameters \todo{Maybe mention HMC methods that could actually do this}. Instead, we pull an Eastern-European trick and Taylor expand theory around $\bar{\vN}$ \todo{we could change this sentence or leave it as is. But I think Taylor was British.}:
      \begin{equation}
        \vt({\bf q},\vN) \simeq \vt({\bf q},\bar{\vN}) + {\sf T} \left(\vN - \bar{\vN} \right), \label{eq:taylor}
      \end{equation}
      where the matrix ${\sf T}$ is the gradient of ${\bf t}$ with respect to $\vN$
      \begin{equation}
        {\sf T}\equiv \left.\frac{d{\bf t}}{d\vN}\right|_{{\bf q},\bar{\vN}}
      \end{equation}
      In other words it contains a full information about how all the correlation functions or power spectra respond to a small change in $N(z)$.

      Substituting Eq. \ref{eq:taylor} into Eq. \ref{eq:like2} results in an integral that is now quadratic in $\vN$ and can be performed analytically. After completing squares, and applying the Woodbury matrix identity, we find:
      \begin{equation}
        \mathcal{L} \propto \left[{\rm det}\left({\sf T}^T {\sf C}_c^{-1} {\sf T} +{\sf P}^{-1}\right)\right]^{-1/2} \exp\left [-\frac{1}{2} ({\bf c}-\vt)^T {\sf C}_M^{-1} ({\bf c}-\vt) \right], \label{eq:main1}
      \end{equation}
      where the  marginalized covariance is
      \begin{equation}
        {\sf C}_M = {\sf C}_c + {\sf T}{\sf P}{\sf T}^T \label{eq:main2}
      \end{equation}
      This is a fascinatingly simple equation and the main result of this paper. This calculation should be understood as follows: for each deviation around $\hat{\vN}$, there is a corresponding deviation around $\vt$. These are directions in space of theory predictions that are perfectly degenerate with the changes in shape of $N(z)$.  Equation \ref{eq:main2}  increases the variance for those linear combinations commensurate with how far ${\sf P}$ allows them to go. The limit ${\sf P}\rightarrow\infty\mathbb{1}$ would correspond to completely projecting out those ``$N(z)$-sensitive'' modes from the data altogether. This is the main result of the paper.

      Both the marginalized covariance ${\sf C}_M$ normalizing prefactor in Eq. \ref{eq:main1} depend in principle on ${\bf q}$ through the parameter dependence of ${\sf T}$. This implies that, in principle, ${\sf T}$ should be re-evaluated at every point in the MCMC chain. Since the calculation of ${\sf T}$ is expensive (by comparison with that of e.g. $\vt$) using standard methods, we will neglect this dependence and evaluate ${\sf T}$ at a fiducial set of parameters. We will however explore the impact of the choice of fiducial parameters on the final results. This should be a sufficiently good approximation for compact likelihoods, where parameter constraints are driven by the differences between $\hat{\bf c}$ and $\vt$. We will therefore pre-compute $T$ once, which then means that for a fixed data covariance matrix ${\sf C}_c$, the normalization term in equation \ref{eq:main1} is an irrelevant constant, and the marginalized covariance in the Eq. \ref{eq:main2} also needs to be evaluated only once.

      The procedure to obtain constraints from this marginalized likelihood would therefore proceed as follows:
      \begin{itemize}
        \item Run inference problem once with fixed redshift distribution and a standard Gaussian likelihood for $\hat{\bf c}$ with covariance ${\sf C}_c$ to determine a sensible fiducial model.
        \item Calculate ${\sf T}$ and ${\sf C}_M$ at this fiducial model.
        \item Re-run inference with a standard Gaussian likelihood using the modified covariance matrix ${\sf C}_M$. This will lead to broadened contours due to the marginalization over redshift distribution uncertainties.
      \end{itemize}
      If needed, one could repeat the inference at a refined fiducial model although in practice we found this to be unnecessary.

      Before moving on, it is worth emphasizing the two distinct approximation used here. The first one is that the theory can validly be Taylor expanded in $\vN-\bar{\vN}$. The second is that the model dependence of ${\sf T}$ can be ignored for the models of interest. We will examine validity of both of those approximations in Section \ref{sec:hsc}.

    \subsection{The prior matrix ${\sf P}$}\label{ssec:theory.prior}
      A key part of this method is the determination of the $N(z)$ prior covariance ${\sf P}$, which governs the amplitude of the uncertainties on $\vN$. As discussed in the previous section, ${\sf P}$ receives two contributions: the covariance associated with the uncertainties in the measured $\hat{\bf N}$, ${\sf C}_N$, and the external prior with covariance ${\sf C}_P$, both combined in an inverse-variance way (see Eq. \ref{eq:priorcov}). We will describe the models used for both contributions in this section.

      \subsubsection{$N(z)$ uncertainties}\label{sssec:theory.prior.cv}
        \begin{figure}[ht]
          \centering  
          \includegraphics[width=1.\textwidth]{./corr_CV_0.png}
          \caption{Correlation matrix associated to the cosmic-variance contribution to the total $N(z)$ covariance for the first tomographic bin of the HSC data. \todo{Can we make these plots pdfs instead?}}\label{fig:CV}
        \end{figure}
        The analysis presented in Section \todo{add ref} will make use of a measured redshift distribution estimated from the COSMOS 30-band catalog \todo{add ref}, as was done in \todo{add cite}. The main sources of uncertainty for this measurement are cosmic variance in the COSMOS field, shot noise and additional uncertainty in the photometric redshifts used to assign galaxies to different bins. Here we will associate each of this sources to a separate covariance that will be added in quadrature to form the final ${\sf C}_N$.

        The contribution from the photometric redshift systematic uncertainties is modelled as \ldots\todo{Not sure how you guys implemented this.}

        The cosmic variance uncertainties are caused by the fluctuations in the matter density traced by galaxies in the particular sky patch. The corresponding covariance would ideally be estimated from simulations including both non-linear gravitational clustering and a realistic model of the galaxy-halo connection for the specific galaxy sample under study. For simplicity in this proof-of-concept analysis we instead use an analytical model for the $N(z)$ covariance. Although less precise than a simulation-based calculation, this approach was shown by \todo{cite sanchez} to provide a reasonable prediction for the redsthift distribution uncertainties.

        The covariance matrix element between $N^\alpha_i$ and $N^\beta_j$ two tomographic bins is given by:
        \begin{equation}\label{eq:nz_cv}
          {\sf C}_{N,{\rm CV}}^{(i\,\alpha),(j\,\beta)}=\frac{N^\alpha_iN^\beta_j}{2\pi^2} \int_0^\infty dk_\parallel \cos(k_\parallel(\chi_\alpha-\chi_\beta))\int_0^\infty dk_\perp k_\perp W_\alpha(k_\parallel,k_\perp)W_\beta(k_\parallel,k_\perp)P_{gg}({\bf k}),
        \end{equation}
        where $\chi_\alpha$ is the radial comoving distance to redshift $z_\alpha$, and we model the galaxy power spectrum using the Kaiser formula \todo{cite} accounting for redshift-space distortions:
        \begin{equation}
          P_{gg}({\bf k},z)=\left(b_g(z)+f(z)\frac{k_\parallel^2}{k^2}\right)^2\,P_{mm}(k,z),
        \end{equation}
        where $f(z)$ is the logarithmic growth rate and $b_g(z)$ is the linear galaxy bias. Following the results of \todo{cite}, appropriate for the magnitude-limited sample studied here, we assume a redshift dependence for $b_g$ given by $b_g(z) = 0.95/D(z)$, with $D(z)$ the linear growth factor. $P_mm(k)$ is the non-linear matter power spectrum, which we model via the \todo{cite} HALOFIT parametrization.

        The overlapping deep HSC patch covering the COSMOS 30-band footprint used to measure the redshift distribution in this work covers a total area of $A_{\rm sky} = 1.7 \ {\rm deg}^2$. Modelling this patch as a disc of radius $\theta_{\rm sky}=0.73^\circ$, the corresponding window function in Eq. \ref{eq:nz_cv} is given by
        \begin{equation}
          W_\alpha(k_\parallel,k_\perp)=j_0(k_\parallel\Delta\chi_\alpha/2)\,\frac{2 J_1(k_\perp \chi_\alpha\theta_{\rm sky})}{k_\perp \chi_\alpha\theta_{\rm sky}},
        \end{equation}
        where $\Delta\chi_\alpha$ is the comoving width of the $\alpha$-th histogram bin, $j_0$ is the zero-th order spherical Bessel function and $J_1$ is the order-1 cylindrical Bessel function.

        We assume that the cosmic covariance between the tomographic bins is negligible, so we estimate the cosmic variance covariance matrix by treating each bin independently. We show the resulting correlation matrix of the COSMOS cosmic variance in Fig. \ref{fig:CV}.
